Energy stable discontinuous Galerkin formulations and discrete differential operators

### 2017-05-06 docs/dg_penalty/revision/main.tex
adding more notes to split form

%\maketitle
\begin{frontmatter}
\title{A short note on the penalty flux parameter for \note{discontinuous Galerkin formulations of first order hyperbolic systems}}

@@

Lemma~\ref{lemma:eig} illustrates the behavior of the \note{spectrum} of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.  However, it is less clear how the \note{spectrum} of $\bm{K}$ behaves for $\tau \approx 1$.  We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.  {\note{Since $\tau$ can be chosen to be any arbitrary non-negative value, we consider three cases: $\tau = 0$, $\tau = 1$, and $\tau \rightarrow \infty$.  The choice of $\tau = 1$ is somewhat arbitrary, and (as mentioned in Section~\ref{sec:scaletau}) it may be beneficial to scale $\tau$ with problem parameters.  However, for the simple model problems considered in this section, taking $\tau = 1$ recovers an upwind flux, and the resulting spectrum of the DG operator is clearly distinct from the cases when $\tau = 0$ or $\tau \rightarrow \infty$.}  }

@@

\begin{figure}
\centering
\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.48\textwidth]{figs/trackedEigs.png}}
\hspace{1em}
\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.435\textwidth]{figs/trackedEigsZoom.png}}
\caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.  Eigenvalues are overlaid on these paths for $\tau = 0$, $\tau = 1$, and $\tau = 4$.  The zoomed in view near the imaginary axis shows the return of spurious modes to the imaginary axis for $\tau$ sufficiently large. }
\label{fig:track1D}
\end{figure}

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigsDiverge3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModesDiverge3.png}}
\caption{Behavior of a mode for the advection equation corresponding to a divergent eigenvalue (indicated by the red star) as $\tau$ increases.  The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.  }
\label{fig:trackmodes1D}
\end{figure}

However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.  Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a $C^0$ continuous function.  These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.  These modes closely resemble back-propagating spurious modes observed for high order $C^0$ finite element methods \cite{ainsworth2014dispersive}, which consist of higher frequency components with sharp peaks.  Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{figs/tauEigs3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{figs/tauModes1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{figs/tauModes2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{figs/tauModes3.png}}
\caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.  For $\tau = 1$ certain eigenvalues have negative real part, and the corresponding eigenmodes are damped.  For $\tau \gg 1$, certain eigenvalues return to the imaginary axis as spurious modes in conforming discretizations. }
\label{fig:spurious}
\end{figure}

Figure~\ref{fig:spurious} also indicates that taking $\tau = 1$ suppresses spurious non-dissipative modes associated with both $\tau = 0$ (central fluxes) and $\tau\rightarrow \infty$ (conforming discretizations).  Let $\lambda^{\tau}_i, \bm{v}^{\tau}_i$ be eigenvalues and eigenvectors of the DG discretization matrix $\bm{K}$ with penalty $\tau$.  The exact solution to the semi-discrete formulation is 
\[
\bm{u}(t) =\bm{V}^{\tau}\LRp{ \sum_j c_j e^{-\lambda_i^{\tau} t} \bm{v}^{\tau}_j}, \qquad \bm{u}(0) = \sum_j c_j\bm{v}^{\tau}_j,
\]
which illustrates how modes $\bm{v}^{\tau}_j$ are damped for $t > 0$ if the corresponding eigenvalue $\lambda^{\tau}_j$ has a negative real part .  We observe that, when spurious modes of the solution for $\tau = 0$ or $\tau \gg 1$ are expanded in the modes for $\tau = 1$
\[
\bm{v}^{\tau}_i = \sum_j c^\tau_j \bm{v}^{1}_j, 
\]
the largest components correspond to eigenvalues with larger negative real parts and heavily damped modes.  As an example, we consider the spurious mode illustrated in Figure~\ref{fig:spurious}.  The absolute values of the coefficients $c^\tau_j$ of this mode are plotted in Figure~\ref{fig:cmodes} for both $\tau = 0$ and $\tau = 100$.  Only the four coefficients larger than $10^{-13}$ are shown.  For comparison, the negative real values of the corresponding eigenvalues $\lambda^1_j$ are also overlaid in Figure~\ref{fig:cmodes}.  The two largest coefficients correspond to eigenvalues with large negative real parts and more strongly damped modes.  The two remaining coefficients correspond to less highly damped eigenvalues with smaller negative real parts, but are an order of magnitude smaller than the two large coefficients.  
\begin{figure}
\centering
\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{figs/ceigs0.png}}
\hspace{1em}
\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{figs/ceigs100.png}}
\caption{Coefficients $c^0_j$ and $c^{100}_j$ of the spurious mode in Figure~\ref{fig:spurious} when expanded in the modes of $\bm{K}$ for $\tau=1$.  Coefficients with magnitude less than $10^{-13}$ are not shown.   The negative real parts of the corresponding eigenvalues ${\rm Re}(\lambda^1_j)$, which determine the magnitude of damping applied to the $j$th component, are overlaid. }
\label{fig:cmodes}
\end{figure}
%\note{Show that both $\tau = 0, \infty$ spurious modes have large damped contributions.}

@@

%\begin{figure}
%\centering
%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs1.png}}
%\hspace{.5em}
%\subfloat{\includegraphics[width=.475\textwidth]{figs/tauEigs2.png}}
%\caption{ }
%\label{fig:trackModesWave}
%\end{figure}

@@

\begin{figure}
\centering
\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{figs/waveEigs1.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{figs/waveEigs2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{figs/waveEigs3.png}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
\label{fig:waveeigs}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{figs/waveEigs2Dbox.png}\label{subfig:big}}
\hspace{2em}
\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{figs/trackedWaveEigs2D.png}\label{subfig:box}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.  Figure~\ref{subfig:box} shows a zoom of the boxed region in Figure~\ref{subfig:big}, with overlaid eigenvalue paths as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
\label{fig:trackEigsWave}
\end{figure}

@@

\begin{figure}
\centering
\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_p3.png}}
\caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
\label{fig:trackModesWave}
\end{figure}

%\begin{figure}
%\centering
%\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u1.png}}
%\hspace{.5em}
%\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u2.png}}
%\hspace{.5em}
%\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{figs/waveEig2D_u3.png}}
%\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
%\label{fig:trackModesWaveU}
%\end{figure}

@@

\begin{figure}
\centering
\subfloat{\includegraphics[width=.475\textwidth]{figs/trackedAdvecEigs.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.465\textwidth]{figs/trackedAdvecEigsZoom1.png}}
\caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
\label{fig:trackModesAdvec}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{figs/advecEig2D_u3.png}}
\caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }
\label{fig:trackModesAdvecU}
\end{figure}

>>>>

%\maketitle
\begin{frontmatter}
\title{On the \note{penalty stabilization mechanism for upwind discontinuous Galerkin formulations of first order hyperbolic systems}}

@@

Lemma~\ref{lemma:eig} illustrates the behavior of the \note{spectrum} of the DG discretization for the asymptotic cases when $\tau = 0$ or $\tau \rightarrow \infty$.  However, it is less clear how the \note{spectrum} of $\bm{K}$ behaves for $\tau \approx 1$.  We rely instead on numerical experiments in one and two space dimensions to illustrate behaviors for the advection and acoustic wave equations.  {\note{Since $\tau$ can be chosen to be any arbitrary non-negative value, we consider three cases: $\tau = 0$, $\tau = 1$, and $\tau \rightarrow \infty$.  As mentioned in Section~\ref{sec:scaletau}, it may be beneficial to scale $\tau$ with problem parameters.  However, for the simple model problems considered in this section, taking $\tau = 1$ recovers an upwind flux, and the resulting spectrum of the DG operator is clearly distinct from the cases when $\tau = 0$ or $\tau \rightarrow \infty$.}  }

@@

\begin{figure}
\centering
\subfloat[Paths of eigenvalues as $\tau$ increases]{\includegraphics[width=.48\textwidth]{trackedEigs.png}}
\hspace{1em}
\subfloat[Zoom of eigenvalues near imaginary axis]{\includegraphics[width=.435\textwidth]{trackedEigsZoom.png}}
\caption{Eigenvalue paths for DG advection with $\tau \in [0,4]$ on a mesh of 8 elements of degree $N=3$.  Eigenvalues are overlaid on these paths for $\tau = 0$, $\tau = 1$, and $\tau = 4$.  The zoomed in view near the imaginary axis shows the return of spurious modes to the imaginary axis for $\tau$ sufficiently large. }
\label{fig:track1D}
\end{figure}

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{tauEigsDiverge3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{tauModesDiverge1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{tauModesDiverge2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{tauModesDiverge3.png}}
\caption{Behavior of a mode for the advection equation corresponding to a divergent eigenvalue (indicated by the red star) as $\tau$ increases.  The order of approximation is $N=3$ on a mesh of 4 elements.} %The displayed eigenfunctions correspond to the highlighted eigenvalue.  }
\label{fig:trackmodes1D}
\end{figure}

However, for sufficiently large $\tau$, a subset of eigenvalues return to the imaginary axis.  Figure~\ref{fig:spurious} illustrates that, as $\tau$ increases, the magnitude of the inter-element jumps present in these modes decreases, and the mode approaches a $C^0$ continuous function.  These eigenvalues which move right to return to the imaginary axis correspond to a second set of \emph{spurious} modes.  These modes closely resemble spurious modes observed for high order $C^0$ finite element methods \cite{boffi2000problem}, which consist of back-propagating high frequency components with sharp peaks \cite{ainsworth2014dispersive}.  Since the real part of these eigenvalues approaches zero as $\tau\rightarrow \infty$, these modes become undamped and persist as spurious solution components in time-domain simulations \cite{hughes2014finite}.  We note that that a similar phenomena has been observed for the solution of steady state problems in \cite{burman2010interior}, where the authors also point out that large penalty parameters result in decreased accuracy of the DG solution.

\begin{figure}
\centering
\subfloat{\includegraphics[width=.32\textwidth]{tauEigs1.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{tauEigs2.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.32\textwidth]{tauEigs3.png}}\\
\subfloat[$\tau = 0$]{\includegraphics[width=.32\textwidth]{tauModes1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$]{\includegraphics[width=.32\textwidth]{tauModes2.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.32\textwidth]{tauModes3.png}}
\caption{Return behavior of DG advection eigenvalues for sufficiently large $\tau$.  For $\tau = 1$ certain eigenvalues have negative real part, and the corresponding eigenmodes are damped.  For $\tau \gg 1$, certain eigenvalues return to the imaginary axis as spurious modes in conforming discretizations. }
\label{fig:spurious}
\end{figure}

Figure~\ref{fig:spurious} also indicates that taking $\tau = 1$ suppresses spurious non-dissipative modes associated with both $\tau = 0$ (central fluxes) and $\tau\rightarrow \infty$ (conforming discretizations).  Let $\lambda^{\tau}_i, \bm{v}^{\tau}_i$ be eigenvalues and eigenvectors of the DG discretization matrix $\bm{K}$ with penalty $\tau$.  The exact solution to the semi-discrete formulation is 
\[
\bm{u}(t) =\bm{V}^{\tau}\LRp{ \sum_j c_j e^{-\lambda_i^{\tau} t} \bm{v}^{\tau}_j}, \qquad \bm{u}(0) = \sum_j c_j\bm{v}^{\tau}_j,
\]
which illustrates how modes $\bm{v}^{\tau}_j$ are damped for $t > 0$ if the corresponding eigenvalue $\lambda^{\tau}_j$ has a negative real part .  We observe that, when spurious modes of the solution for $\tau = 0$ or $\tau \gg 1$ are expanded in the modes for $\tau = 1$
\[
\bm{v}^{\tau}_i = \sum_j c^\tau_j \bm{v}^{1}_j, 
\]
the largest components correspond to eigenvalues with larger negative real parts and heavily damped modes.  As an example, we consider the spurious mode illustrated in Figure~\ref{fig:spurious}.  The absolute values of the coefficients $c^\tau_j$ of this mode are plotted in Figure~\ref{fig:cmodes} for both $\tau = 0$ and $\tau = 100$.  Only the four coefficients larger than $10^{-13}$ are shown.  For comparison, the negative real values of the corresponding eigenvalues $\lambda^1_j$ are also overlaid in Figure~\ref{fig:cmodes}.  The two largest coefficients correspond to eigenvalues with large negative real parts and more strongly damped modes.  The two remaining coefficients correspond to less highly damped eigenvalues with smaller negative real parts, but are an order of magnitude smaller than the two large coefficients.  
\begin{figure}
\centering
\subfloat[$c_j^0$ (spurious modes for  $\tau = 0$)]{\includegraphics[width=.45\textwidth]{ceigs0.png}}
\hspace{1em}
\subfloat[$c_j^{100}$ (spurious modes for $\tau = 100$)]{\includegraphics[width=.45\textwidth]{ceigs100.png}}
\caption{Coefficients $c^0_j$ and $c^{100}_j$ of the spurious mode in Figure~\ref{fig:spurious} when expanded in the modes of $\bm{K}$ for $\tau=1$.  Coefficients with magnitude less than $10^{-13}$ are not shown.   The negative real parts of the corresponding eigenvalues ${\rm Re}(\lambda^1_j)$, which determine the magnitude of damping applied to the $j$th component, are overlaid. }
\label{fig:cmodes}
\end{figure}
%\note{Show that both $\tau = 0, \infty$ spurious modes have large damped contributions.}

@@

%\begin{figure}
%\centering
%\subfloat{\includegraphics[width=.475\textwidth]{tauEigs1.png}}
%\hspace{.5em}
%\subfloat{\includegraphics[width=.475\textwidth]{tauEigs2.png}}
%\caption{ }
%\label{fig:trackModesWave}
%\end{figure}

@@

\begin{figure}
\centering
\subfloat[$\tau = 1$]{\includegraphics[width=.3\textwidth]{waveEigs1.png}}
\hspace{.5em}
\subfloat[$\tau = 10$]{\includegraphics[width=.3\textwidth]{waveEigs2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$]{\includegraphics[width=.3\textwidth]{waveEigs3.png}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions. Note the changing scale of the real axis.}
\label{fig:waveeigs}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 0,1, 50$, zoomed in view]{\includegraphics[height=.3\textheight]{waveEigs2Dbox.png}\label{subfig:big}}
\hspace{2em}
\subfloat[$\tau= 0,1,50$, returning eigenvalue paths]{\includegraphics[height=.3\textheight]{trackedWaveEigs2D.png}\label{subfig:box}}
\caption{Behavior of eigenvalues for the acoustic wave equation in two dimensions.  Figure~\ref{subfig:box} shows a zoom of the boxed region in Figure~\ref{subfig:big}, with overlaid eigenvalue paths as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
\label{fig:trackEigsWave}
\end{figure}

@@

\begin{figure}
\centering
\subfloat[$\tau = .1$, $\lambda = -0.2379 + 8.7528i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$, $\lambda = -1.0145 + 8.9270i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$, $\lambda = -0.0437 + 7.6167i$]{\includegraphics[width=.3\textwidth]{waveEig2D_p3.png}}
\caption{Behavior of the real part of the pressure component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
\label{fig:trackModesWave}
\end{figure}

%\begin{figure}
%\centering
%\subfloat[$\tau = .1$, $\lambda = -0.6009 +11.3209i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u1.png}}
%\hspace{.5em}
%\subfloat[$\tau = 1$, $\lambda =  -1.9601 + 9.5629i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u2.png}}
%\hspace{.5em}
%\subfloat[$\tau = 100$, $\lambda = -0.0634 + 7.7670i$]{\includegraphics[width=.3\textwidth]{waveEig2D_u3.png}}
%\caption{Behavior of the real part of the velocity component of spurious eigenmodes for the acoustic wave equation in two dimensions.  }
%\label{fig:trackModesWaveU}
%\end{figure}

@@

\begin{figure}
\centering
\subfloat{\includegraphics[width=.475\textwidth]{trackedAdvecEigs.png}}
\hspace{.5em}
\subfloat{\includegraphics[width=.465\textwidth]{trackedAdvecEigsZoom1.png}}
\caption{Eigenvalue paths for the DG discretization of advection as $\tau$ increases.  Divergent eigenvalues in the far left half plane are not shown.  }
\label{fig:trackModesAdvec}
\end{figure}

\begin{figure}
\centering
\subfloat[$\tau = 0$, $\lambda = 0 + 7.9246i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u1.png}}
\hspace{.5em}
\subfloat[$\tau = 1$, $\lambda =  -0.8278 + 6.7831i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u2.png}}
\hspace{.5em}
\subfloat[$\tau = 100$, $\lambda = -0.0239 + 5.9282i$]{\includegraphics[width=.3\textwidth]{advecEig2D_u3.png}}
\caption{Behavior of the real part of spurious eigenmodes for the advection equation with $\beta = (1,0)$. }
\label{fig:trackModesAdvecU}
\end{figure}

### 2017-05-06 docs/splitform/main.tex
adding more notes to split form

\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[titletoc,toc,title]{appendix}

@@

\newcommand{\note}[1]{{\color{blue}{#1}}}
\newcommand{\remark}[1]{\textbf{\color{red}#1}}

@@

\date{}
\author{Jesse Chan}
\title{Split form DG methods}

@@

\maketitle
%\tableofcontents

\section{Notes}

Integration by parts gives
\[
\sum_{D^k} \LRp{\Grad u, \bm{v}}_{\LK} = \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{u,\bm{v}\cdot\bm{n}}_{\LdK}.
\]
We replace the values of ${u}$ with $\avg{u}$ on each element boundary to define the global DG gradient operator $\Grad_h$ 
\[
\sum_{D^k} \LRp{\Grad_h u, \bm{v}}_{\LK} \coloneqq \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{\avg{u},\bm{v}\cdot\bm{n}}_{\LdK}.
\]
Integrating by parts again and the introduction of the lift operator shows that
\[
\Grad_h u = \Grad u + \frac{1}{2}L\LRp{\jump{u}\bm{n}}
\]
where $L$ is the lift operator.

The DG divergence operator is similarly defined as
\[
\sum_{D^k} \LRp{\Grad_h \cdot \bm{u}, v}_{\LK} \coloneqq \sum_{D^k} \LRp{-\bm{u}, \Grad v}_{\LK} + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\LdK}.
\]
and
\[
\Grad_h \cdot u = \Grad\cdot u + \frac{1}{2}L\LRp{\jump{\bm{u}}\cdot\bm{n}}
\]
and it can be shown that, for a periodic mesh,
\[
\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}.
\]
We will incorporate boundary conditions in a stable way in the following sections.

@@

\bibliographystyle{unsrt}
\bibliography{dgpenalty}

>>>>

\usepackage{fullpage}
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem*{remark}{Remark}
%\usepackage{thmtools}
%\declaretheorem[style=definition,qed=$\blacksquare$,numberwithin=chapter]{definition}

\usepackage[titletoc,toc,title]{appendix}

@@

\newcommand{\note}[1]{{\color{blue}{#1}}}

@@

\date{}
\author{Jesse Chan}
\title{Energy stable discontinuous Galerkin formulations and discrete differential operators}

@@

\maketitle

\begin{abstract}
We show that, for non-linear hyperbolic conservation laws which admit a skew-symmetric splitting, energy stable discontinuous Galerkin (DG) methods can be constructed in a straightforward manner based on discrete DG gradient and divergence operators.  In particular, we show that these formulations remain energy stable for both curvilinear geometries and inexact quadrature.  Examples of energy stable formulations are given for variable advection, Burgers' equation, and a Burgers'-like system, and it is shown that the construction of DG methods based on discrete differential operators recover known entropy-conservative fluxes.  
\end{abstract}

\section{Introduction}

\cite{feng2016discontinuous}

Lifting operators  and discrete gradients \cite{bassi1997high, di2011mathematical}

\section{Discrete differential operators}

We assume that the domain $\Omega$ is decomposed into non-overlapping elements $D^k$.  We define the mesh $\Omega_h = \cup D^k$ and the corresponding global approximation space $V_h(\Omega_h) = \bigoplus V_h\LRp{D^k}$, where $V_h\LRp{D^k}$ is the approximation space over a single patch.  Furthermore, we introduce the jump and average of discontinuous functions across element interfaces.  Let $D^{k,+}$ denote the neighboring element of across a face $f$ of $D^k$, and let $u^+,u^-$ denote the values of $u$ on $D^{k,+}$ and $D^k$, respectively.  The jump of $u$ across $f$ is then defined as
\[
\jump{u} = u^+ - u^-, \qquad \avg{u} = \frac{u^+ + u^-}{2}.
\]
The jump and average of vector fields are defined component-wise using the jumps and averages of components.

\subsection{Discrete inner products}

In order to define discrete differential operators, we first introduce the $L^2$ inner product on $V_h\LRp{\Oh}$
\[
\LRp{u,v}_{\Omega} = \sum_{D^k} \int_{D^k} uv \diff{x} = \sum_{D^k} \int_{\widehat{D}} uv J \diff{\widehat{x}}.  
\]
In practice, these integrals are computed using quadrature, such that 
\[
\int_{\widehat{D}}uv = \sum_{i=1}^{N_q} u(\bm{x}_i)v(\bm{x}_i) w_i,
\]
where $N_q$ is the number of quadrature points.  The only assumptions we make upon this quadrature is that it is sufficiently accurate such that 
\begin{enumerate}
\item The quadrature induces an equivalent inner product over the reference element $\widehat{D}$
\item Integration by parts holds (with respect to the reference coordinates $\widehat{x}$).
\end{enumerate}

We also introduce the $L^2$ projection $\Pi_N: L^2\LRp{\Omega}\rightarrow P^N$ such that 
\[
\LRp{\Pi_N u,v}_{\Omega} = \LRp{u,v}_{\Omega}.
\]

\subsection{Discrete derivatives}
We introduce two discrete DG derivatives in this section.  The first is the implicit definition used in \cite{di2011mathematical}

\begin{definition}
The discontinuous Galerkin differentiation operator ${D}^i: L^2\LRp{\Omega}\rightarrow V_h$ is defined implicitly as follows:
\[
\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{-u,\pd{v}{\bm{x}_i}}_{\LK} + \LRa{\avg{u},v\bm{n}_i}_{\LdK}}.
\]
\end{definition}

We note that, for $u,v \in V_h$, integration by parts yields an equivalent definition of $D^i$ found in \cite{hesthaven2004high,Warburton20063205}
\[
\LRp{{D}^i u,v}_{\Oh} = \sum_{D^k} \LRp{\LRp{\pd{u}{\bm{x}_i},v}_{\LK} + \frac{1}{2}\LRa{\jump{u},v\bm{n}_i}_{\LdK}}.
\]
However, for $u \not\in V_h$, these two definitions are not equivalent.  This motivates the definition of a \textit{discrete} DG differentiation operator:
\begin{definition}
The discrete discontinuous Galerkin differentiation operator ${D}^i_h: L^2\LRp{\Omega} \rightarrow V_h$ is defined as $D^i_h = D^i \Pi_N$.
\end{definition}

The main property of

\section{Notes}
Begin with the gradient.  Multiplication by $\bm{v}$ and integration by parts gives
\[
\sum_{D^k} \LRp{\Grad u, \bm{v}}_{\LK} = \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{u,\bm{v}\cdot\bm{n}}_{\LdK}.
\]
We replace the values of ${u}$ with $\avg{u}$ on each element boundary to define the global DG gradient operator $\Grad_h$ 
\[
\sum_{D^k} \LRp{\Grad_h u, \bm{v}}_{\LK} \coloneqq \sum_{D^k} \LRp{-u, \Grad \cdot \bm{v}}_{\LK} + \LRa{\avg{u},\bm{v}\cdot\bm{n}}_{\LdK}.
\]
Integrating by parts again and the introduction of the lift operator shows that
\[
\Grad_h u = \Grad u + \frac{1}{2}L\LRp{\jump{u}\bm{n}}
\]
where $L$ is the lift operator.

The DG divergence operator is similarly defined as
\[
\sum_{D^k} \LRp{\Grad_h \cdot \bm{u}, v}_{\LK} \coloneqq \sum_{D^k} \LRp{-\bm{u}, \Grad v}_{\LK} + \LRa{\avg{\bm{u}}\cdot\bm{n},v}_{\LdK}.
\]
and
\[
\Grad_h \cdot u = \Grad\cdot u + \frac{1}{2}L\LRp{\jump{\bm{u}}\cdot\bm{n}}
\]
and it can be shown that 
\[
\LRp{\Grad_h u, \bm{v}} = \LRp{-u, \Grad_h \cdot \bm{v}}.
\]
We will incorporate boundary conditions in a stable way in the following sections.

@@

\bibliographystyle{unsrt}
\bibliography{dg}
